---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dev = "ragg_png",
  dpi = 200
)
```

# psyarxivdbr

psyarxivdbr is a poorly named R package that provides access to [PsyArXiv](https://osf.io/preprints/psyarxiv) preprint metadata within R. The data is sourced from the [Open Science Framework](https://osf.io/) API, and is updated monthly.

## Install

Install psyarxivdbr from [GitHub](https://github.com/mvuorre/psyarxivdbr):

```r
# install.packages("pak")
pak::pkg_install("mvuorre/psyarxivdbr")
```

## Use

```{r}
#| results: hide
#| message: false

library(psyarxivdbr)
library(tidyverse) # Data wrangling
library(knitr) # Table display
```

The package provides a single table of preprint metadata called `preprints`. Here's what the data look like (see `?preprints`):

```{r}
glimpse(preprints)
```

### Example: Preprints over time

As an example, we tabulate the number of preprints created per year.

```{r}
#| label: tbl-year
#| tbl-cap: Yearly new PsyArXiv preprints.

preprints |>
  mutate(
    year = floor_date(date_created, unit = "year") |> str_sub(1, 4)
  ) |>
  count(year) |>
  pivot_wider(names_from = year, values_from = n) |>
  kable()
```
